---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Setting global options
```{r, echo = TRUE, warning = FALSE}
library(knitr)
library(kableExtra)
opts_chunk$set(echo = T, dpi = 200)
options(knitr.table.format = "html") 
```

## Define a display function to help us show the table more elegantly
```{r}
display <- function(t){
  kable(t, align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
}
```

## Loading and preprocessing the data

### 1. Unzip activity.zip
```{r unzip}
zipfile <- "activity.zip"
unzip(zipfile, exdir = ".")
datapath <- "activity.csv"
```

### 2. Load the data
```{r loaddata, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(printr)
df <- read_csv(datapath)
display(head(df, 10))
```

## What is mean total number of steps taken per day?

### 1. Calculate the total number of steps per day
```{r}
daydf <- df %>%
  group_by(date) %>%
  summarise(steps = sum(steps, na.rm = T))
display(head(daydf, 10))
```

### 2. Make a histogram for the total steps per day
```{r, warning = FALSE}
library(ggplot2)
ggplot(daydf, aes(steps)) +
  geom_histogram(bins = 20) +
  ggtitle("Histogram for Total Steps per Day") +
  theme(plot.title = element_text(hjust = 0.5))
```

### 3. Calculate the mean and median of the total steps per day
```{r}
result <- daydf %>%
  summarise(mean = mean(steps), median = median(steps))
display(result)
```

## What is the average daily activity pattern?

### 1. Calculate the average steps for each time interval
```{r}
timedf <- df %>%
  group_by(interval) %>%
  summarise(steps = mean(steps, na.rm = T)) %>%
  mutate(minutes = interval %/% 100 * 60 + interval %% 100) %>%
  select(-steps, steps)
display(head(timedf, 10))
```

### 2. Make a time series plot
```{r}
# Define a function that can transform interval into a label to help us plot
interval_to_label <- function(x){
  paste(formatC(x %/% 60, width = 2, flag = "0"), 
        formatC(x %% 60, width = 2, flag = "0"), 
        sep = ":")
}
ggplot(timedf, aes(x = minutes, y = steps)) +
  geom_line() +
  ggtitle("Time Series Plot of Average Steps") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("5-minute interval") +
  scale_x_continuous(breaks = seq(0, 1440, 180), labels = interval_to_label)
```

### 3. Find the interval that contains the maximum average steps
```{r, message = FALSE}
result <- timedf %>%
  top_n(n = 1)
display(result)
```

## Imputing missing values

### 1. Calculate the total number of missing values in the dataset
```{r}
result <- summarise_all(df, function(x){sum(is.na(x))})
display(result)
```

Now, we can see that only the steps column has missing values. Therefore, there are **2304 rows with NAs** in the dataset.

### 2. Fill in all the missing values with mean for 5-minute interval
```{r}
fixeddf <- df %>%
  group_by(date, interval) %>%
  mutate(steps = if_else(is.na(steps), timedf$steps[which(timedf$interval == interval)], as.double(steps)))
display(head(fixeddf, 10))
```

### 3. Calculate the total number of steps per day
```{r}
fixeddaydf <- fixeddf %>%
  group_by(date) %>%
  summarise(steps = sum(steps, na.rm = T))
display(head(fixeddaydf, 10))
```

### 4. Make a histogram of total steps per day
```{r, warning = FALSE}
ggplot(fixeddaydf, aes(steps)) +
  geom_histogram(bins = 20) +
  ggtitle("Histogram for Total Steps per Day (Fixed)") +
  theme(plot.title = element_text(hjust = 0.5))
```

### 3. Calculate the mean and median of the total steps per day
```{r}
result <- fixeddaydf %>%
  summarise(mean = mean(steps), median = median(steps))
display(result)
```

The mean and median are **larger than** the estimates from the first part of the assignment. The total daily numbers of steps of the days with missing values become **larger** after the missing data is imputed, while others remain the same.

## Are there differences in activity patterns between weekdays and weekends?

### 1. Create a factor variable with two levels
```{r}
fixeddf <- fixeddf %>%
  mutate(day = 
           if_else(weekdays(date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))
display(head(fixeddf, 10))
```

### 2. Calculate the total average steps by group of interval and day
```{r}
fixedtimedf <- fixeddf %>%
  group_by(interval, day) %>%
  summarise(steps = mean(steps, na.rm = T)) %>%
  mutate(minutes = interval %/% 100 * 60 + interval %% 100) %>%
  select(-steps, steps)
display(head(fixedtimedf, 10))
```

### 3. Create a time series plot comparing weekends and weekdays

```{r}
ggplot(fixedtimedf, aes(x = minutes, y = steps, color = day)) +
  geom_line() +
  facet_grid(day ~.) +
  ggtitle("Time Series Plot of Average Steps on Different Days (Fixed)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("5-minute interval") +
  scale_x_continuous(breaks = seq(0, 1440, 180), labels = interval_to_label)
```

