---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Setting global options
```{r, echo = TRUE}
library(knitr)
opts_chunk$set(echo = T, dpi = 200)
```

## Loading and preprocessing the data

### 1. Unzip activity.zip
```{r unzip}
zipfile <- "activity.zip"
unzip(zipfile, exdir = ".")
datapath <- "activity.csv"
x <- 3
```

### 2. Load the data
```{r loaddata, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(printr)
df <- read_csv(datapath)
head(df)
```

## What is mean total number of steps taken per day?

### 1. Calculate the total number of steps per day
```{r}
daydf <- df %>%
  group_by(date) %>%
  summarise(steps = sum(steps, na.rm = T))
head(daydf)
```

### 2. Make a histogram for the total steps per day
```{r}
library(ggplot2)
ggplot(daydf, aes(steps)) +
  geom_histogram(bins = 20) +
  ggtitle("Histogram for Total Steps per Day") +
  theme(plot.title = element_text(hjust = 0.5))
```

### 3. Calculate the mean and median of the total steps per day
```{r}
daydf %>%
  summarise(mean = mean(steps), median = median(steps))
```

## What is the average daily activity pattern?

### 1. Calculate the average steps for each time interval
```{r}
timedf <- df %>%
  group_by(interval) %>%
  summarise(steps = mean(steps, na.rm = T))
head(timedf)
```

### 2. Make a time series plot
```{r}
ggplot(timedf, aes(x = interval, y = steps)) +
  geom_line() +
  ggtitle("Time Series Plot of Average Steps") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("5-minute interval") +
  scale_x_continuous(breaks = seq(0, 2400, 240), labels = function(x){
    formatC(x, width = 4, flag = "0", big.interval = 2, big.mark = ":")})
```

### 3. Find the interval that contains the max average steps
```{r, message = FALSE}
timedf %>% 
  top_n(n = 1)
```

## Imputing missing values

### 1. Calculate the total number of missing values in the dataset
```{r}
summarise_all(df, function(x){sum(is.na(x))})
```

Now, we can see that only the steps column has missing values. Therefore, there are **2304 rows with NAs** in the dataset.

### 2. Fill in all the missing values with mean for 5-minute interval
```{r}
fixeddf <- df %>%
  mutate(steps = 
           if_else(is.na(steps), timedf$steps[which(timedf$interval == interval)], as.double(steps)))
head(fixeddf, 10)
```

### 3. Calculate the total number of steps per day
```{r}
fixeddaydf <- fixeddf %>%
  group_by(date) %>%
  summarise(steps = sum(steps, na.rm = T))
head(fixeddaydf, 10)
```

### 4. Make a histogram of total steps per day
```{r}
ggplot(fixeddaydf, aes(steps)) +
  geom_histogram(bins = 20) +
  ggtitle("Histogram for Total Steps per Day (Fixed)") +
  theme(plot.title = element_text(hjust = 0.5))
```

### 3. Calculate the mean and median of the total steps per day
```{r}
fixeddaydf %>%
  summarise(mean = mean(steps), median = median(steps))
```

The mean and median are **different from** the estimates from the first part of the assignment. The total daily numbers of steps of the days with missing values become **larger** after the missing data is imputed, while others remain the same.

## Are there differences in activity patterns between weekdays and weekends?
